---
title: "class09_mini-project"
author: 'Rui Huang (PID: A15606522)'
date: "10/26/2021"
output:
  pdf_document: default
  html_document: default
---

```{r}
fna.data<-"WisconsinCancer.csv"
wisc.df <- read.csv(fna.data, row.names=1)
```

```{r}
wisc.data <- wisc.df[,-1]
diagnosis <- as.factor(wisc.df[,1])
```

#Q1: there are 569 observations
```{r}
dim(wisc.data)
```

#Q2: 212 observations of malignant diagnosis
```{r}
table(diagnosis)
```

#Q3:there is 10 variables are sufficed with mean
```{r}
length(grep("mean", colnames(wisc.df)))
```

```{r}
colMeans(wisc.data)
apply(wisc.data,2,sd)
wisc.pr<- prcomp(wisc.data,scale=T)
summary(wisc.pr)
```

#Q4:the original variance captured by the PC1 has a proportion of 0.4427.

#Q5: 3 principle components are required.

#Q6: 7 principle components are required.

```{r}
biplot(wisc.pr)
```
#Q7: the plot is crowded as there are way too many data on it, it is hard to understand since there is too much information with many overlapping scripts.

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
```

#Q8:the shown y-axis window range is different, and also the distribution of the dots are different.
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3],col=diagnosis, xlab="PC1", ylab="PC3")
```

```{r}
df<- as.data.frame(wisc.pr$x)
df$diagnosis<- diagnosis
library(ggplot2)
ggplot(df)+aes(PC1,PC2, col=diagnosis)+geom_point()
```

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```
```{r}
pve<-pr.var/sum(pr.var)
plot(pve,xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1), type="o")
```


```{r}
barplot(pve,ylab="Percent of Variance Explained", names.arg=paste0("PC",1:length(pve)),las=2,axes=F)
axis(2, at=pve, labels=round(pve,2)*100)
```
```{r}
library(factoextra)
fviz_eig(wisc.pr, addlables=T)
```

#Q9: it is -0.2608538
```{r}
wisc.pr$rotation["concave.points_mean",1]
```

#Q10: the minimum is 5
```{r}
var<-summary(wisc.pr)
var$importance[3,]>=0.8
```
```{r}
data.scaled<- scale(wisc.data)
data.dist<-dist(data.scaled)
wisc.hclust<- hclust(data.dist)
```

#Q11:at the height of 19, the clustering model has 4 clusters.
```{r}
plot(wisc.hclust)
abline(h=19,col="red",lty=2)
```
```{r}
wisc.hclust.clusters<- cutree(wisc.hclust, k=4)
table(wisc.hclust.clusters,diagnosis)
```

#Q12:cutting them into 2 clusters is a better way.
```{r}
wisc.hclust.clusters1<- cutree(wisc.hclust, k=2)
table(wisc.hclust.clusters1,diagnosis)
```

#Q13:My favorite method is "word.D2", because the clusters branching corresponds to greater height range, so that it is better presented and organized and it is easier to see the precise height of the branching clusters.
```{r}
hc.single<-hclust(data.dist, method="single")
hc.complete<-hclust(data.dist, method="complete")
hc.average<-hclust(data.dist, method="average")
hc.ward<-hclust(data.dist, method="ward.D2")
plot(hc.single)
plot(hc.complete)
plot(hc.average)
plot(hc.ward)
```
#kmeans
```{r}
wisc1<-scale(wisc.data, center=T,scale=T)
wisc.km<- kmeans(wisc1,centers=2, nstart=20)
table(wisc.km$cluster, diagnosis)
```
```{r}
table(wisc.hclust.clusters,wisc.km$cluster)
```
#Q14: k-means separate the two diagnosis well, its separation of 2 diagnosis is as good as hclust result.

#5.Combing methods
```{r}
summary(wisc.pr)
wisc.pc.hclust<-hclust(dist(wisc.pr$x[,1:3]),
                       method="ward.D2")
```

```{r}
plot(wisc.pc.hclust)
abline(h=60,col="red")
```

```{r}
summary(wisc.pr)
wisc.pr.hclust<-hclust(dist(wisc.pr$x[,1:7]),
                       method="ward.D2")
wisc.pr.hclust.cluster <- cutree(wisc.pr.hclust, k=2)

```

```{r}
grps<- cutree(wisc.pr.hclust, k=2)
table(grps)
```

#cross table compare of diagnosis and my cluster group
```{r}
table(diagnosis,grps)
```
```{r}
plot(wisc.pr$x[,1:2], col=grps)
```
```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
```

#Q15: the separation of diagnosis outcomes is good,as the cluster 1 mainly corresponds  to malignant, and cluster 2 corresponds to benign diagnosis.
```{r}
table(wisc.pr.hclust.cluster,diagnosis)
```
#Q16:Both the k-means and the hierarchial clustering separate the diagnosis well, and both of them are eqaully good in terms of separating the diagnosis.
```{r}
table(wisc.km$cluster, diagnosis)
```

```{r}
table(wisc.hclust.clusters, diagnosis)
```
#sensitivity
```{r}
188/(188+24)
175/(175+37)
165/(165+47)
```

#specificity
```{r}
329/(329+28)
343/(343+14)
343/(343+14)
```

#Q17:wisc.pr.hclust.cluster' result has better sensitivity, k means and wisc.hclust.clusters' result have better specificity.

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```


```{r}
g <- as.factor(grps)
g <- relevel(g,2)
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")

```
#Q18: we should prioritize patient 2.
```{r}
sessionInfo()
```




