---
title: "class17:covid-19 vaccination rate mini project"
author: 'Rui Huang (PID: A15606522)'
date: "11/23/2021"
output:
  pdf_document:
    latex_engine: xelatex
---


#background

```{r}
vax <- read.csv("covid19vac.csv")
head(vax)
```
#Q1:the column 9 "persons_fully_vaccinated"

```{r}
head(vax[,9])
```


#Q2:second column "zip_code_tabulation_area"

#Q3:2021-01-05

#Q4:2021-01-05
```{r}
skimr::skim(vax)
```

#Q5:9 numeric columns

#Q6: 8256 missing values
```{r}
sum( is.na(vax$persons_fully_vaccinated) )
```


#Q7: 10% of value is missing
```{r}
8256/81144
```


#Q8: the zip code area has no data to report.
```{r}
library(lubridate)
```



#Q: how many days since last entry
```{r}
vax$as_of_date <- ymd(vax$as_of_date)
today()-vax$as_of_date[nrow(vax)]
```

#Q9: 315 days.
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

#Q10: 46 unique dates.
```{r}
length(unique(vax$as_of_date))
```

```{r}
library(zipcodeR)
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
table(vax$county)
```

subset with base R
```{r}
inds <- vax$county == "San Diego"
head(vax[inds,])
```

use **dplyr** and its filter()
```{r}
library(dplyr)
sd <- filter(vax, county=="San Diego")
head(sd)
```

```{r}
nrow(sd)
```

#Q11: 106 distinct zip codes
```{r}
length(unique(sd$age12_plus_population))
```

#Q12: it is 92154
```{r}
ind <- which.max(sd$age12_plus_population)
sd[ind,]
```

```{r}
filter(sd, zip_code_tabulation_area == "92037")[1,]
```


#Q13: 67.27%
```{r}
sd.now <- filter(sd, as_of_date == "2021-11-09")

mean(sd.now$percent_of_population_fully_vaccinated, na.rm=T)
```

```{r}
summary(sd.now$percent_of_population_fully_vaccinated)
```

#Q14:
```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```

```{r}
library(ggplot2)
ggplot(sd.now)+aes(percent_of_population_fully_vaccinated)+geom_histogram(bins=15)

```
92037-la jolla/ucsd

```{r}
 filter(sd.now, zip_code_tabulation_area == "92037")
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

#Q15
```{r}
ggplot(ucsd)+
  aes(as_of_date, percent_of_population_fully_vaccinated)+
  geom_point()+
  geom_line(group=1)+
  ylim(c(0,1))+
  labs(x="Date", y="Percent Vaccinated")
```

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")
mean(vax.36$percent_of_population_fully_vaccinated, na.rm=T)
```

#Q16:
```{r}
ggplot(ucsd)+
  aes(as_of_date, percent_of_population_fully_vaccinated)+
  geom_point()+
  geom_line(group=1)+
  geom_hline(yintercept=0.6629812, col="red", linetype=2)
```

#Q17:
```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```




#Q18: 

```{r}
ggplot(vax.36)+
  aes(percent_of_population_fully_vaccinated)+
  geom_histogram()
```

#Q19: 92040 is below the averageof vax.36, but the 92109 is above the average of vax.36.
```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```


```{r}
vax.36.all <- filter(vax, age5_plus_population>36144)
nrow(vax.36.all)
```

```{r}
length(unique(vax.36.all$zip_code_tabulation_area))
```

#Q20:
```{r}
ggplot(vax.36.all)+ 
  aes(as_of_date, percent_of_population_fully_vaccinated, group=zip_code_tabulation_area)+
  geom_line(alpha=0.2, color="blue")+
  ylim(c(0,1))+
  labs(x="Date",y="Percent Vaccinated",title="Vaccination rate accross California",subtitle="Only areas with a population above 36k are shown")+
  geom_hline(yintercept= 0.66,col="red", linetype=2)
```

